Subject-level developmental relationships

### 2024-08-29 1_get_data.qmd
extra changes from morning

@@

```{r}
d_trial <- d_trial_rt_unfiltered |>
  filter(!(is.na(short_window_prop_data) & is.na(long_window_prop_data)), 
         !(short_window_prop_data < .5 & long_window_prop_data < .5), 
         age >= 12) |>
  mutate(across(contains("window"), ~ ifelse(is.na(.x), NA, .x))) |>
  mutate(short_window_accuracy = ifelse(short_window_prop_data >= .5, short_window_accuracy, NA),
         long_window_accuracy = ifelse(long_window_prop_data >= .5, long_window_accuracy, NA),
         rt = ifelse(short_window_prop_data >= .5, rt, NA),
         short_window_elogit = ifelse(short_window_prop_data >= .5, short_window_elogit, NA), 
         long_window_elogit = ifelse(long_window_prop_data >= .5, long_window_elogit, NA))

>>>>

Let's look at the RT distribution.

```{r}
ggplot(d_trial_rt_unfiltered, aes(x = rt)) + 
  geom_histogram() + 
  geom_vline(xintercept = 367, col = "red", lty =2 )
```

@@

Let's also remove RTs < 367 ms. These are unlikely to be real.

```{r}
d_trial <- d_trial_rt_unfiltered |>
  filter(!(is.na(short_window_prop_data) & is.na(long_window_prop_data)), 
         !(short_window_prop_data < .5 & long_window_prop_data < .5), 
         age >= 12) |>
  mutate(across(contains("window"), ~ ifelse(is.na(.x), NA, .x))) |>
  mutate(short_window_accuracy = ifelse(short_window_prop_data >= .5, 
                                        short_window_accuracy, NA),
         long_window_accuracy = ifelse(long_window_prop_data >= .5, 
                                       long_window_accuracy, NA),
         rt = ifelse(short_window_prop_data >= .5, rt, NA),
         short_window_elogit = ifelse(short_window_prop_data >= .5, 
                                      short_window_elogit, NA), 
         long_window_elogit = ifelse(long_window_prop_data >= .5, 
                                     long_window_elogit, NA), 
         rt = ifelse(rt < 367, NA, rt), 
         log_rt = ifelse(rt < 367, NA, log_rt))

### 2024-08-29 3_developmental_relationships.qmd
extra changes from morning

```{r}
d_cdi <- cdi_data |>
  select(administration_id, native_language, measure, language, 
         rawscore) |>
  filter(language == "English (American)", native_language == "eng") |>
  pivot_wider(names_from = "measure", values_from = "rawscore", 
              values_fill = NA, values_fn = mean)
```

@@

```{r}
GGally::ggpairs(select(d_sub_cdi, log_age, rt,  
                       long_window_accuracy, prod, comp), 
        progress = FALSE, lower = list(continuous = GGally::wrap("points", alpha = 0.03)))

@@

<!-- So these are somewhat related to one another. Let's take a deeper look.  -->

<!-- ```{r} -->
<!-- ggplot(acc_rt, aes(x = log_rt, y = elogit, col = log_age)) + -->
<!--   geom_point(alpha = .3) + -->
<!--   geom_smooth(method = "lm") + -->
<!--   geom_hline(yintercept = .5, lty = 2) +  -->
<!--   xlab("Mean log RT") +  -->
<!--   ylab("Mean elogit(accuracy)") -->
<!-- ``` -->

<!-- Broken out by age group.  -->

<!-- ```{r} -->
<!-- ggplot(acc_rt, aes(x = log_rt, y = elogit)) + -->
<!--   geom_point(alpha = .3) + -->
<!--   geom_smooth(method = "lm") + -->
<!--   geom_hline(yintercept = .5, lty = 2) + -->
<!--   facet_wrap(~age_group_years) +  -->
<!--   xlab("Mean log RT") +  -->
<!--   ylab("Mean elogit accuracy") -->

@@

@@

d_sub_cdi_mat <- d_sub_cdi_mat |>
  filter(complete.cases(d_sub_cdi_mat)) |>
  as.matrix()

>>>>

```{r}
d_cdi <- cdi_data |>
  select(administration_id, native_language, measure, language, 
         CDI_percent) |>
  filter(language == "English (American)", native_language == "eng") |>
  pivot_wider(names_from = "measure", values_from = "CDI_percent", 
              values_fill = NA, values_fn = mean)
```

@@

```{r}
GGally::ggpairs(select(d_sub_cdi, log_age, rt,  
                       long_window_accuracy, short_window_accuracy, prod, comp), 
        progress = FALSE, lower = list(continuous = GGally::wrap("points", alpha = 0.03)))

@@

# Individual relationships

```{r}
m_rt <- mean(d_sub_cdi$rt, na.rm=TRUE)
sd_rt <- sd(d_sub_cdi$rt, na.rm=TRUE)
d_sub_cdi$factor_rt <- cut(d_sub_cdi$rt, 
                           breaks = quantile(d_sub_cdi$rt, 
                                             probs = c(0, .25, .5, .75, 1), 
                                             na.rm= TRUE))

ggplot(filter(d_sub_cdi, !is.na(rt), !is.na(factor_rt)),
       aes(x = log_age, 
                      y = long_window_accuracy, 
                      col = factor_rt)) +
  geom_point(alpha = .3) +
  geom_smooth() +
  geom_hline(yintercept = .5, lty = 2)+
  viridis::scale_color_viridis(discrete = TRUE, 
                               name = "RT quantile")

Broken out by age group.

```{r}

d_sub_cdi$age_group <- cut(d_sub_cdi$log_age, 
                           breaks = log(c(12,24,36,48,60)), 
                           labels = c("12-24","24-36","36-48","48-60"), 
                           include_lowest = TRUE)

ggplot(d_sub_cdi, 
       aes(x = rt, 
                      y = long_window_accuracy, 
                      col = age_group)) +
  geom_point(alpha = .3) +
  geom_smooth() +
  geom_hline(yintercept = .5, lty = 2)+
  viridis::scale_color_viridis(discrete = TRUE, 
                               name = "age group")
```

@@

Make into matrices.

d_sub_cdi_mat <- d_sub_cdi_mat |>
  filter(complete.cases(d_sub_cdi_mat)) |>
  as.matrix() 
```

Factor analysis

```{r}
# scree(d_sub_cdi_mat, pc=FALSE)  # Use pc=FALSE for factor analysis

# install.packages("psych")
mod1 <- fa(d_sub_cdi_mat, 1)
mod2 <- fa(d_sub_cdi_mat, 2)

@@

```{r}
